Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
corol 4.8

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
  \item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    then $\judgeDreduce{s_2}{s_4}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_4$
  \item If $\judgeDreduce{s_2}{s_4}$
    then $\judgeSreduce{C}{s_1}{s_3}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_3$
\end{enumerate}
{\color{red}
  Jeremy's attempt to reformulate in a more standard way that
  remains equivalent:
\begin{enumerate}
\item If $\judgeSreduce{C}{s_1}{s_3}$ then
  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
\item If $\judgeDreduce{s_2}{s_4}$ then
  $\judgeSreduce{C}{s_1}{s_3}$,
  $\judgeDreduceTrans{s_2}{s_5}$, and
   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
}
Or diagrammatically, 
  \[\begin{array}{clclc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) follow from the definition of $s \approx s$.
  Part (2) and (3) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

@@

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:SD-equiv}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
  Proving that the left-hand side implies the right-hand side is
  trivial.  To prove $\judgeDeval{e}{o}$ implies
  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
  catch it up in \ineffCEKD\ by taking one or more steps in
  \ineffCEKD. As we just took a positive number of steps, we are
  closer to the halting state.
\end{proof}

>>>>

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
  \item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    then $\judgeDreduce{s_2}{s_4}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_4$
  \item If $\judgeDreduce{s_2}{s_4}$
    then $\judgeSreduce{C}{s_1}{s_3}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_3$
\end{enumerate}
{\color{red}
  Jeremy's attempt to reformulate in a more standard way that
  remains equivalent:
\begin{enumerate}
\item If $\judgeSreduce{C}{s_1}{s_3}$ then
  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
\item If $\judgeDreduce{s_2}{s_4}$ then
  $\judgeSreduce{C}{s_1}{s_3}$,
  $\judgeDreduceTrans{s_4}{s_5}$, and
   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
}
Or diagrammatically, 
  \[\begin{array}{clclc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) follow from the definition of $s \approx s$.
  Part (2) and (3) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

@@

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:SD-equiv}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
	By the definitions of $eval$, we need to show 
	\[
	\judgeSreduceTrans{C}{\sOOinspect{e}{\emptyset}{wrap(\kOOmt)}}{
		\sOOhalt{o}
	}
	\;\text{if and only if}\;
	\judgeSreduceTrans{C}{\sOOinspect{e}{\emptyset}{\kOOmt}}{
		\sOOhalt{o}
	}
	\]
	Obviously, the initial states are related in the bisimulation relation.
	\[
	\sOOinspect{e}{\emptyset}{wrap(\kOOmt)}
	\approx
	\sOOinspect{e}{\emptyset}{\kOOmt}
	\]
	So we can generalize our goal, and prove that if $s_1 \approx s_2$
	\[
	\judgeSreduceTrans{C}{s_1}{\sOOhalt{o}}
	\;\text{if and only if}\;
	\judgeSreduceTrans{C}{s_2}{\sOOhalt{o}}
	\]
	Let us prove left to right first by induction on 
	$\judgeSreduceTrans{C}{s_1}{\sOOhalt{o}}$.

	\begin{itemize}
		\item 
		If $s_1$ takes zero step to the halting state, it must be halting.
		By Lemma~\ref{thm:SD-bisim}, $s_2$ is halting as well, so we are done.
		\item 
		If $\judgeSreduce{C}{s_1}{s_3}$ and 
		$\judgeSreduceTrans{C}{s_3}{\sOOhalt{o}}$,
		by Lemma~\ref{thm:SD-bisim}, there exists a $s_4$ such that 
		$\judgeDreduceTrans{s_2}{s_4}$ and $s_3 \approx s_4$. So we can apply 
		our induction hypothesis.
	\end{itemize}

	The other direction is similar.
%  Proving that the left-hand side implies the right-hand side is
%  trivial.  To prove $\judgeDeval{e}{o}$ implies
%  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
%  catch it up in \ineffCEKD\ by taking one or more steps in
%  \ineffCEKD. As we just took a positive number of steps, we are
%  closer to the halting state.
\end{proof}
